{\tt APFELgrid}: a fast interface to hadronic observables for fits of parton densities

### 2016-05-03 paper.tex
conclusions pass

%\section{The {\tt APFELgrid} package}
% **** Not sure if this sort of section would be neccesary ***
%The {\tt APFELgrid} plugin generates {\tt FK} tables of the form
%expressed in Eq.~(\ref{eq:FKTable}) from weight grids supplied by {\tt
%  APPLgrid} files. It is supplied as a set of {\tt C++} routines designed to
%  supplement the PDF evolution library {\tt APFEL} with {\tt FK} table generation
%  capabilities. The plugin has a very light interface for performing the {\tt FK} combination,
%  and relies on the native {\tt APFEL} API for setting the parameters of the PDF evolution,
%  including the structure of the initial scale $x$ grid.
%*** Would now discuss e.g SIMD implementation etc in advance of looking at the results in next section ***
\section{Performance benchmarks}
\label{sec:benchmark}

@@

In this work we have presented {\tt APFELgrid}, an efficient
convolution interface to hadronic observables for fits of parton
densities. We have demonstrated that employing the so-called {\tt
  FastKernel} method, it is possible to convert an {\tt APPLgrid}
weight table into a derived format, referred to as {\tt FKtable},
which includes PDF and $\alpha_s$ evolution as provided by {\tt
  APFEL}.  Such a procedure allows one to obtain a very
computationally-efficient expression for the calculation of hadronic
cross sections, in terms of only the initial-scale PDFs and summing
only over the initial scale interpolating grid and the incoming parton
flavors. In addition, the simple structure of the resulting
convolution makes {\tt FKtables} particularly suitable to make
efficient use of acceleration and parallelisation techniques like {\tt
  SIMD} and {\tt OpenMP}.

We have shown that the numerical convolution of an {\tt APFELgrid}
tables is typically much faster than the corresponding {\tt
  APPLgrid}. In particular, the {\tt FastKernel} method provides a
great performance improvement with respect to {\tt APPLgrid} which is
further enhanced by the employment of any of the acceleration
techniques available in {\tt APFELgrid}. Finally, while the size of an
{\tt FKtable} is often larger than the corresponding {\tt APFELgrid}
file, we have shown that the resident set size memory required by {\tt
  APFELgrid} is typically smaller than {\tt APPLgrid} by at least
75\%.

The substantial speed improvement with respect to {\tt APPLgrid} in
association with the reduction of the resident set size memory, makes
the {\tt APFELgrid} code a valuable tool for modern PDF fits which
usually incluse thousands of hadronic data points and that thus
heavily rely on fast interfaces. We stress that {\tt APFELgrid} is not
a replacement for {\tt APPLgrid}, on which it depends, but it is
rather a specialised tool conceived to ease the inclusion of hadronic
data into PDF fits. In fact, the price of a better performance is the
fact that, by themselves, {\tt FKtables} cannot perform any parameter
variation other than that of PDFs. For example, variations of
$\alpha_s$ can be performed only by re-computing the {\tt FKtables}.

The {\tt APFELgrid} code is publicly available from the webpage:
\begin{center}
\url{https://github.com/nhartland/APFELgrid}
\end{center}
where intructions to install and run codes are also provided.

>>>>

\section{Performance benchmarks}
\label{sec:benchmark}

@@

In this work we have demonstrated that by employing the so-called {\tt
  FastKernel} method, it is possible to convert an {\tt APPLgrid}
weight table into a derived format, referred to as {\tt FK} table,
which includes PDF and $\alpha_s$ evolution. This procedure has been 
implemented in the {\tt APFELgrid} package, supplied as a set of {\tt C++} routines designed to
supplement the PDF evolution library {\tt APFEL} with {\tt FK} table generation
capabilities. 
%The plugin has a very light interface for performing the {\tt FK} combination,
%  and relies on the native {\tt APFEL} API for setting the parameters of the PDF evolution,
%  including the structure of the initial scale $x$ grid.
The {\tt APFELgrid} package allows one to obtain a 
computationally efficient expression for the calculation of hadronic
cross sections, in terms of only the initial-scale PDFs and summing
only over the initial scale interpolating $x$-grid and the incoming parton
flavors. Additionally the simple structure of the resulting
convolution makes {\tt FK} tables particularly suitable to make
efficient use of computational tools such as {\tt
  SIMD} and {\tt OpenMP}.

We have shown that in several practical examples the numerical evaluation of an {\tt FK}
product is much faster than the corresponding {\tt APPLgrid} product, even in the case where
neither SIMD or multi-threading are applied. {\tt FK} tables are supplied in a simple plain-text 
format in order to simplify the construction of user interfaces, and therefore typically are often larger 
than the corresponding {\tt APFELgrid} file. However we have shown that the in-memory resident set size occupied by {\tt
  FK} tables is typically smaller than that requred by {\tt APPLgrid} by at least 75\%.

The substantial speed improvement of {\tt FK} tables with respect to {\tt APPLgrid} in
association with the reduction in memory footprint, makes
the {\tt APFELgrid} code a valuable tool for modern PDF fits including large collider datasets and thus
rely heavily upon fast calculations.

%We stress that {\tt APFELgrid} is not
%a replacement for {\tt APPLgrid}, on which it depends, but it is
%rather a specialised tool conceived to ease the inclusion of hadronic
%data into PDF fits. In fact, the price of a better performance is the
%fact that, by themselves, {\tt FKtables} cannot perform any parameter
%variation other than that of PDFs. For example, variations of
%$\alpha_s$ can be performed only by re-computing the {\tt FKtables}.

The {\tt APFELgrid} package and associated documentation are publicly available from the webpage:
\begin{center}
\url{https://github.com/nhartland/APFELgrid}
\end{center}
